### 2021-07-13 src/paper/model/matching.tex
Add formal description of matching.

Recurrent contacts are described by two components: 1) A set of time invariant groups, such as school classes. The groups can be sampled from empirical data or created by
randomly matching simulated individuals into groups. 2) A deterministic or random function that takes the value 0 (non-participating) and 1
(participating) and can depend on the weekday, date and health state. This can be used
to model vacations, weekends or symptomatic people who stay home (see section
\ref{sec:endogenous_contact_reductions} for details).

The matching process for recurrent contacts is then extremely simple: On each simulated
day, every person who does not stay home meets all other group members who do not stay
home. The assumption that all group members have contacts with all other group members
is not fully realistic, but a good approximation to reality, especially in
light of the suspected role of aerosol transmission for Covid-19 \citep{Morawska2020,
Anderson2020}. Alternatively, the infection probability of recurrent contact types can
be interpreted as being the product of a true infection probability and the probability
that an actual contact takes place.

The matching in non-recurrent contact models is more difficult and implemented in a two
stage sampling procedure to allow for assortative matching. Currently most contact
models are assortative with respect to age (it is more likely to meet people from the
same age group) and county (it is more likely to meet people from the same county) but
in principle any set of discrete variables can be used. This set of variables that
influence matching probabilities introduce a discrete partition of the population into
groups. The first stage of the two stage sampling process samples on the group level.
The second stage on the individual level.

The algorithm works as follows: First we randomly draw a contact type and individual.
For each contact of the drawn contact type that person has, we first draw the group of
the other person (first stage).
Next, we calculate the probability to be drawn for each member of the group, based on
the number of remaining contacts, i.e. people who have more remaining contacts are drawn
with a higher probability. The probabilities have to be re-calculated each time because
with each
matched contact, the number of remaining contacts changes. We then draw the other
individual, determine whether an infection takes place and if so update the health state
of the newly infected person. Finally, we reduce the number of remaining contacts of the
two matched individuals by one.\comment[id=J]{Add formulae here.}

The recalculation of matching probabilities in the second stage is computationally
intensive because it requires summing up all remaining contacts in that group. Using a
two stage sampling process where the first stage probabilities remain constant over time
makes the matching computationally much more tractable because the number of
computations only increases quadratically in the size of the second group and not
quadratically in the size of the entire simulated population.

>>>>

%% Recurrent contacts ============================================================
Recurrent contacts are described by two components: 1) A set of time invariant groups,
such as school classes or groups of co-workers. Those groups are generated once at the
beginning of the simulation. The groups can be sampled from empirical data or created by
randomly matching simulated individuals into groups. 2) A deterministic or random
function that takes the value 0 (non-participating) and 1 (participating) and can depend
on the weekday, date and health state. This can be used to model vacations, weekends or
symptomatic people who stay home (see section \ref{sec:policies} for details).

Given those two components, the disease transmission for recurrent contacts is extremely
simple: On each simulated day, every person who does not stay home meets all other group
members who do not stay home. If there is a contact between individual $i$ who is
infected with virus variant $v$ and infectious and individual $j$ who is in age group
$a$ and susceptible, then $j$ becomes infected with the following probability

\begin{equation}
    \label{eq:infection_probability_formula}
    P(infection) = \beta_c \cdot s_{c, t} \cdot \sigma_v \cdot \zeta_a
\end{equation}

where $beta_c$ denotes the base infection probability of contact type $c$,
$s_{c, t}$ is a seasonality factor between zero and one that depends on the contact
type $c$ and time $t$ (see Equation~\ref{eq:seasonality}), $\sigma_v$ is the
infectiousness factor of virus variant $v$ and $\zeta_a$ is an age dependent
susceptibility factor.

The assumption that all group members have contacts with all other group members
is not fully realistic, but a good approximation to reality, especially in
light of the suspected role of aerosol transmission for Covid-19 \citep{Morawska2020,
Anderson2020}. Alternatively, the infection probability of recurrent contact types can
be interpreted as being the product of a true infection probability and the probability
that an actual contact takes place.

%% Non recurrent contacts

The matching of non-recurrent contact types is more difficult because the contact
network is resampled randomly every day. Moreover, it needs to allow for
assortative matching. In our application, all
random contacts are assortative with respect to age group $a$ (it is usually more likely
to meet
people from the same age group) and county (it is more likely to meet people from the
same county) but in principle any set of discrete variables can be used. This set of
variables that influence matching probabilities introduce a discrete partition of the
population into groups.

Below we first describe one iteration of a simplified matching algorithm that
illustrates what we want to achieve. In practice, we approximate the result of this
matching algorithm by a two stage sampling procedure that is computationally more
efficient. The matching is done for each non-recurrent contact type $c$. The following
step is repeated until no individual has unmatched contacts left. Let $z$ be an iteration
counter for the matching algorithm and $i$ denote the individual whose unmatched
contacts we are trying to match.

Let $K_{z,i,c}$ denote the number of unmatched contacts of individual $i$ of contact type
$c$ before iteration $z$ is completed. Note that $K_{z,i,c} \leq n_{ic}$ which is the
total number of contacts individual $i$ has of type $c$.

Let $a_i$ denote $i$'s age group and $county_i$ her county of residence.

We first draw individual $j$ from the distribution defined by probability mass function
$F_{z}$ over individuals $j \neq i$ in the synthetic population where the probability
$f_{zj}$ is calculated as follows:

\begin{equation}
    \label{eq:combined_matching_probabilities}
    f_{zj} = \underbrace{\alpha_{c,a_i,a_j, county_i, county_j}\vphantom{\frac{bla}{\sum_i^k}}}_{\text{Group Probability}} \cdot \underbrace{\frac{K_{z,j,c}}{\sum_{l=1}^N K_{z,l,c} \cdot \mathbb{I}_{county_l=county_j \land a_l = a_j}}}_{\text{Individual probability}}
\end{equation}

We then draw an individual $j$. If one of the two participants is susceptible and the
other one is infectious, we sample whether an infection takes place. The success
probability for this event is calculated as in
Equation~\ref{eq:infection_probability_formula}. Finally we update the remaining
numbers of unmatched contacts by setting:

\begin{align}
    K_{z+1,i,c} = K_{z, i, c} - 1
    K_{z+1, j, c} = K_{z, j, c} - 1
\end{align}

The runtime of this algorithm scales roughly cubic in the number $N$ of simulated
individuals. This is because the number of iterations is proportional to $N$, in each
iteration we have evaluate Equation~\ref{eq:combined_matching_probabilities} $N$ times
and each evaluation of that equation entails a sum over $N$ individuals.

This makes it prohibitively expensive. We
therefore replace the above algorithm by a two stage sampling procedure, where we
first sample the group from which individual $j$ will be drawn according to the
group probabilities defined in Equation~\ref{eq:combined_matching_probabilities}.
Next we sample an individual from this group with the Individual probabilities defined
in Equation~\ref{eq:combined_matching_probabilities}.

Thus, while the calculation of any given second stage probability entails exactly the
same number of calculations as before we do not have to calculate a second stage
probability for all simulated individuals but only for those who are members of
the group that was sampled in the first stage.

It is easy to see that ex-ante the probability of being sampled are identical between
the two stage sampling and the one stage sampling. The only drawback is
that towards the end of the matching process it becomes possible to sample a group in
which no unmatched contacts are left. In our empirical application this happens
extremely rarely. This is so for two reasons: Firstly, the first stage sampling
probabilities have been estimated from the same dataset as the numbers of contacts so
there cannot be any mismatches such as for example a group that has a low probability
of being sampled in the first stage but where all members have many contacts.
Secondly, the group sizes are relatively large and we go over individuals in random
order. Therefore, groups where no unmatched contacts remain only occur very late in the
matching process.\footnote{If unmatched contacts were a concern one could simply
use the fast two stage sampling process for a first pass over contacts and then
match all remaining contacts with the slow algorithm.}
